---
title: "HW4_Sarah_Zhou"
author: "Sarah Zhou"
month: "11/29/2015"
output: html_document
runtime: shiny
---

###Loading packages:
```{r}
library(rvest)
library(lubridate)
library(stringr)
library(dplyr)
library(ggplot2)
```

###Part 1: Harvesting the data

**Harvesting Season 2014-2015 data**

Reading the web page

```{r}
page <- read_html("http://www.goduke.com/SportSelect.dbml?SPSID=22726&SPID=1845&DB_OEM_ID=4200&Q_SEASON=2014") 
```

Harvesting the data using the SelectorGadget tool, a javascript bookmarklet that allows us to figure out what css selector each variable is on the page, which we can then pass into the html_nodes function.

I used str-trim to trim the unwanted characters.

I also used tail(-1) to delete the first entry for some of the variables that included the header. The extra header would've made the number of rows unequal.

For the season column, I inserted the string "14-15" into the rep function to repeat it "length(month)" number of times so it would have the same number of rows as the rest of the variables.

```{r}
month <- page %>%
  html_nodes(".date_nowrap") %>%
  html_text() %>%
  str_trim() 
  
season <- rep("14-15",length(month))

opponent <- page %>%
  html_nodes(".opponent") %>%
  html_text() %>%
  str_trim() %>%
  tail(-1) 

location <- page %>%
  html_nodes(".location") %>%
  html_text() %>%
  str_trim() %>%
  tail(-1) 

dukerank <- page %>%
  html_nodes(".team-rank") %>%
  html_text() %>%
  str_trim() %>%
  tail(-1)

opprank <- page %>%
  html_nodes(".opponent-rank") %>%
  html_text() %>%
  str_trim() %>%
  tail(-1)

result <- page %>%
  html_nodes(".results") %>%
  html_text() %>%
  str_trim() %>%
  tail(-1)
```

Storing all the variables into a data frame

```{r}
blue_devils_1415 <- data.frame(season, month, opponent, location, result, dukerank, opprank)
```

**Cleaning up the data**

Convert both dukerank and opprank variable into the character type and then to numeric type so they are numeric values

```{r}
blue_devils_1415$dukerank = as.numeric(as.character(blue_devils_1415$dukerank))
blue_devils_1415$opprank = as.numeric(as.character(blue_devils_1415$opprank))
```

Convert the result variable to the character type so that I can alter it using a function in the stringr package

```{r}
blue_devils_1415$result = as.character(blue_devils_1415$result)
```

Used the grepl function to clean up the result variable. The result variables contain a letter indicating Win (W) or Loss (L) plus a comma and the score of both teams (eg. "W, 34-9"). However, I want to make this variable categorical with two categories (Win and Loss) so I'm not interested in the score, and I also want to indicate "Win" and "Loss" instead of just a letter to make it easier to read. I use the grepl function which is a boolean function that checks if a string is a substring of a string value, returning True if it is or False if it isn't. Using the grepl function and ifelse statements, I mutated the variable so that if "W" is in the value, I replaced it with "Win", and if "L" is in the value, I replaced it with "Loss", thus getting rid of the scores, making the variable easier to read, and overall making it a more useful categorical variable as there are now only two levels to work with. 

```{r}
blue_devils_1415 <- blue_devils_1415 %>%
  mutate(result = ifelse(grepl("W",result), "Win",
                         ifelse(grepl("L",result),"Loss",
                                NA
                         )))
```

Again, I used ifelse statements and the grepl function to mutate the month variable so that instead of displaying the entire date, it only displays the month, making it a more useful categorical variable as well as it would reduce the number of levels to work with to 6.

```{r}
blue_devils_1415 <- blue_devils_1415 %>%
  mutate(month = ifelse(grepl("Jan",month), "Jan",
                       ifelse(grepl("Feb",month),"Feb",
                              ifelse(grepl("Mar",month),"Mar",
                                     ifelse(grepl("Oct",month),"Oct",
                                            ifelse(grepl("Nov",month),"Nov",
                                                   ifelse(grepl("Dec",month),"Dec",
                                                          NA)))))) )
```

To make the location variable a more useful categorical variable, I mutated it so that the value would be "home" if the original value was "Durham, N.C." and "Away" if it wasn't, reducing the number of levels to two. 

```{r}
blue_devils_1415 <- blue_devils_1415 %>%
  mutate(location = ifelse(location == "Durham, N.C.", "Home",
                           "Away"))

blue_devils_1415
```

**Function for harvesting the data and cleaning up data:**

Each function contains a parameter for the HTML link to the page and the seasonNumbers ("13-14", "12-13", etc), which is used to fill the season column.

```{r}
#Harvesting data

seasons <- function(seasonNumbers = NULL, HTML = NULL) {
  
page <- read_html(HTML) 
  
month <- page %>%
  html_nodes(".date_nowrap") %>%
  html_text() %>%
  str_trim() 
  
season <- rep(seasonNumbers,length(month))

opponent <- page %>%
  html_nodes(".opponent") %>%
  html_text() %>%
  str_trim() %>%
  tail(-1) %>%
  str_replace("&amp;#39;","'")

location <- page %>%
  html_nodes(".location") %>%
  html_text() %>%
  str_trim() %>%
  tail(-1) %>%
  str_replace("at ","")

dukerank <- page %>%
  html_nodes(".team-rank") %>%
  html_text() %>%
  str_trim() %>%
  tail(-1)

opprank <- page %>%
  html_nodes(".opponent-rank") %>%
  html_text() %>%
  str_trim() %>%
  tail(-1)

result <- page %>%
  html_nodes(".results") %>%
  html_text() %>%
  str_trim() %>%
  tail(-1)

blue_devils_data <- data.frame(season, month, opponent, location, result, dukerank, opprank)
 
#Cleaning up the data

blue_devils_data$dukerank = as.numeric(as.character(blue_devils_data$dukerank))
blue_devils_data$opprank = as.numeric(as.character(blue_devils_data$opprank))

blue_devils_data$result = as.character(blue_devils_data$result)

blue_devils_data <- blue_devils_data %>%
  mutate(result = ifelse(grepl("W",result), "Win",
                         ifelse(grepl("L",result),"Loss",
                                NA
                         )))

blue_devils_data <- blue_devils_data %>%
  mutate(month = ifelse(grepl("Jan",month), "Jan",
                       ifelse(grepl("Feb",month),"Feb",
                              ifelse(grepl("Mar",month),"Mar",
                                     ifelse(grepl("Oct",month),"Oct",
                                            ifelse(grepl("Nov",month),"Nov",
                                                   ifelse(grepl("Dec",month),"Dec",
                                                          NA)))))) )

blue_devils_data <- blue_devils_data %>%
  mutate(location = ifelse(location == "Durham, N.C.", "Home",
                           "Away"))


}
```

**Calling the function for each season**

```{r}
blue_devils_1314 <- seasons("13-14", "http://www.goduke.com/SportSelect.dbml?SPSID=22726&SPID=1845&DB_OEM_ID=4200&Q_SEASON=2013")
blue_devils_1314

blue_devils_1213 <- seasons("12-13", "http://www.goduke.com/SportSelect.dbml?SPSID=22726&SPID=1845&DB_OEM_ID=4200&Q_SEASON=2012")
blue_devils_1213

blue_devils_1112 <- seasons("11-12", "http://www.goduke.com/SportSelect.dbml?SPSID=22726&SPID=1845&DB_OEM_ID=4200&Q_SEASON=2011")
blue_devils_1112

blue_devils_1011 <- seasons("10-11", "http://www.goduke.com/SportSelect.dbml?SPSID=22726&SPID=1845&DB_OEM_ID=4200&Q_SEASON=2010")
blue_devils_1011
```

**Stacking data sets together and writing it as .csv file**

```{r}
basketball <- rbind(blue_devils_1415,blue_devils_1314,blue_devils_1213,blue_devils_1112,blue_devils_1011)

write.csv(basketball,file="bluedevilsbasketball.csv")
```

###Part 2: Create an interactive visualization of the data

My shiny app allows the user to choose variables on a drop-down menu for two plots: a categorical and numerical one. 

For the categorical one, the user can choose an explanatory categorical variable from a drop-down menu and a responding categorical variable (fill) from another drop-down menu. Using ggplot, the result will be a bar plot with the explanatory variable graphed on the x-axis, the responding variable as the fill, and the count graphed on the y-axis. 

For the numerical one, the user can choose an explanatory numerical variable from a drop-down menu, a responding categorical variable from another drop-down menu, and a categorical variable as the color. Using ggplot, the result will be a scatterplot with the explanatory variable graphed on the x-axis, the responding variable on the y-axis, and the dots colored by the categorical variable. 

```{r}
shinyUI(fluidPage(
  titlePanel("Categorical and Numerical Analysis on Duke Basketball Data"),
  sidebarLayout(
    sidebarPanel(
      
      selectInput("expvar", 
                  "Explanatory Variable (Categorical):",
                  choices = c("result", "season", "location", "month"), selected = "result"),
      
      selectInput("color", 
                  "Responding Variable (Categorical):",
                  choices = c("result","season", "location","month"), selected = "season"),
      
      hr(),
      
      selectInput("numvar1",
                  "Explanatory Variable (Numerical):",
                  choices = c("dukerank","opprank"), selected = "dukerank"),
      
      selectInput("numvar2",
                  "Responding Variable (Numerical):",
                  choices = c("dukerank","opprank"), selected = "opprank"),
      
      selectInput("color2", 
                  "Categorical Variable:",
                  choices = c("result","season", "location","month"), selected = "result")
      
    ),
    
    mainPanel(
      
      plotOutput('categoricalPlot'),
      
      hr(),
      
      plotOutput("numericalPlot")
    )
  )
))

output$categoricalPlot <- renderPlot({
  
  basketball <- basketball %>%
    filter_(!is.na(input$expvar)) %>%
    filter_(!is.na(input$color))
  
  ggplot(data = basketball, aes_string(x = input$expvar, fill = input$color)) +
    geom_bar() +
    ggtitle(paste0("Categorical Plot: ", input$expvar," vs. ",input$color))
  
})

output$numericalPlot <- renderPlot({
  
  basketball <- basketball %>%
    filter(!is.na(input$numvar1)) %>%
    filter(!is.na(input$numvar2))
  
  ggplot(data = basketball, aes_string(x = input$numvar1, y = input$numvar2, color = input$color2)) +
    geom_point() +
    ggtitle(paste0("Numerical Plot: ", input$numvar1," vs. ",input$numvar2))
})
```

